---
title: "Lab 10"
author: "Skyler Moomey"
date: "`r Sys.Date()`"
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

### Working Directory

```{r}
getwd()
```

## Task 2

### mymaxlik()

Line A creates a new matrix, called $z$, which performs the binary operation denoted by lfun using the $i^{th}$ term of $x$ as the first operand and the $j^{th}$ term of param as the second. Thus, the $i,j$ index of $z$ is equal to $x$[i] lfun param[j].

Line B sets $i$ equal to the largest index of the column sums of $z$ where the maximum column sum was found.

### Maximum Likelihood of 8 Binomial Experiments, n=20

$$
\begin{equation*}
  \begin{split}
  L &= P(y_1)...P(y_8)\\
    &=\left ( {20 \choose y_1}p^{y_1}q^{20-y_1} \right)...\left ( {20 \choose y_8}p^{y_8}q^{20-y_8} \right)\\
    &={20 \choose y_1}...{20 \choose y_8}p^{y_1+...+y_8}q^{20 - y_1-...-y_8}\\
    &= \prod_{i=1}^{n}{20 \choose y_i}p^{y_i}(1-p)^{20-y_i}
  \end{split}
\end{equation*}
$$

### R function for this

```{r, eval=FALSE}
dbinom(y1,  20, prob)*dbinom(y2, 20, prob)*dbinom(y3, 20, prob)*dbinom(y4, 20,   prob)*dbinom(y5, 20,  prob)*dbinom(y6, 20, prob)*dbinom(y7, 20,  prob)*dbinom(y8, 20, prob)
```

### Plot

```{r}
mymaxlik=function(lfun,x,param,...){
# how many param values are there?
np=length(param)
# outer -- notice the order, x then param
# this produces a matrix -- try outer(1:4,5:10,function(x,y) paste(x,y,sep=" "))   to understand
z=outer(x,param,lfun)
# z is a matrix where each x,param is replaced with the function evaluated at those values
y=apply(z,2,sum)

# y is a vector made up of the column sums
# Each y is the log lik for a new parameter value
plot(param,y,col="Blue",type="l",lwd=2,...)
# which gives the index for the value of y == max.
# there could be a max between two values of the parameter, therefore 2 indices
# the first max will take the larger indice
i=max(which(y==max(y)))
abline(v=param[i],lwd=2,col="Red")

# plots a nice point where the max lik is
points(param[i],y[i],pch=19,cex=1.5,col="Black")
axis(3,param[i],round(param[i],2))
#check slopes. If it is a max the slope shoud change sign from + to 
# We should get three + and two -vs
ifelse(i-3>=1 & i+2<=np, slope<-(y[(i-2):(i+2)]-y[(i-3):(i+1)])/(param[(i-2):(i+2)]-param[(i-3):(i+1)]),slope<-"NA")
return(list(i=i,parami=param[i],yi=y[i],slope=slope))
}
logbin=function(x,param) log(dbinom(x,prob=param,size=10))
mymaxlik(x=c(3,3,4,3,4,5,5,4), param=seq(0,1, length=1000), lfun="logbin", xlab=expression(pi),main="Binomial",cex.main=2)
```



